---
title: "JEDTHESIS"
output: html_document
---
```{r loading}
pkg <- c("GPArotation", "mvtnorm", "polycor", "psych", "readxl", "scales", "xtable", "ggplot2", "dplyr", "knitr", "kableExtra", "formattable", "MASS", "magrittr", "lmtest", "descr", "lavaan", "devtools", "semPlot", "semTools", "nonnest2", "psycho", "corrplot")
has   <- pkg %in% rownames(installed.packages())
if(any(!has)) install.packages(pkg[!has])

df <- read_excel("/Users/jed/Downloads/school/DATA 793 - THESIS/ESS2-9.xlsx", sheet="ESS2-9")
df <- data.frame(df)
df2 <- read_excel("/Users/jed/Downloads/school/DATA 793 - THESIS/ESS2-9.xlsx", sheet="2004")
df2 <- data.frame(df)
df6 <- read_excel("/Users/jed/Downloads/school/DATA 793 - THESIS/ESS2-9.xlsx", sheet="2012")
df6 <- data.frame(df)
df8 <- read_excel("/Users/jed/Downloads/school/DATA 793 - THESIS/ESS2-9.xlsx", sheet="2016")
df8 <- data.frame(df)
df9 <- read_excel("/Users/jed/Downloads/school/DATA 793 - THESIS/ESS2-9.xlsx", sheet="2018")
df9 <- data.frame(df)
#touristnumbers
#2004 - 360,400
#2012 - 672,900
#2016 - 1,792,200
#2018 - 2,315,925
```

```{r}
df$imsmetn <- ifelse(df$imsmetn  == "7" | df$imsmetn  == "8" | df$imsmetn  == "9", NA, df$imsmetn)
df$imdfetn <- ifelse(df$imdfetn  == "7" | df$imdfetn  == "8" | df$imdfetn  == "9", NA, df$imdfetn)
df$impcntr <- ifelse(df$impcntr  == "7" | df$impcntr  == "8" | df$impcntr  == "9", NA, df$impcntr)

#immigration recodes
df <- df %>%
  mutate(imsmetn = 5 - imsmetn,
         impcntr = 5 - impcntr,
         imdfetn = 5 - imdfetn)

df$imsmetn <- ifelse(df$imsmetn == "1", 1, df$imsmetn)
df$imsmetn <- ifelse(df$imsmetn == "3" | df$imsmetn == "2", 2, df$imsmetn)
df$imsmetn <- ifelse(df$imsmetn == "4", 3, df$imsmetn)

df$imdfetn <- ifelse(df$imdfetn == "1", 1, df$imdfetn)
df$imdfetn <- ifelse(df$imdfetn  == "3" | df$imdfetn == "2", 2, df$imdfetn)
df$imdfetn <- ifelse(df$imdfetn == "4", 3, df$imdfetn)

df$impcntr <- ifelse(df$impcntr == "1", 3, df$impcntr)
df$impcntr <- ifelse(df$impcntr == "3" |df$impcntr == "2", 2, df$impcntr)
df$impcntr <- ifelse(df$impcntr == "4", 3, df$impcntr)


df$imbgeco <- ifelse(df$imbgeco  == "77" | df$imbgeco  == "88" | df$imbgeco  == "99", NA, df$imbgeco)
df$imbgeco <- ifelse(df$imbgeco == "7" | df$imbgeco == "8" | df$imbgeco == "9" | df$imbgeco == "10", 3, df$imbgeco)
df$imbgeco <- ifelse(df$imbgeco == "4" | df$imbgeco == "5" | df$imbgeco == "6", 2, df$imbgeco)
df$imbgeco <- ifelse(df$imbgeco == "0" | df$imbgeco == "1" | df$imbgeco == "2" | df$imbgeco == "3", 1, df$imbgeco)

df$imueclt <- ifelse(df$imueclt  == "77" | df$imueclt  == "88" | df$imueclt  == "99", NA, df$imueclt)
df$imueclt <- ifelse(df$imueclt == "7" | df$imueclt == "8" | df$imueclt == "9" | df$imueclt == "10", 3, df$imueclt)
df$imueclt <- ifelse(df$imueclt == "4" | df$imueclt == "5" | df$imueclt == "6", 2, df$imueclt)
df$imueclt <- ifelse(df$imueclt == "0" | df$imueclt == "1" | df$imueclt == "2" | df$imueclt == "3", 1, df$imueclt)

df$imwbcnt <- ifelse(df$imwbcnt  == "77" | df$imwbcnt  == "88" | df$imwbcnt  == "99", NA, df$imwbcnt)
df$imwbcnt <- ifelse(df$imwbcnt == "7" | df$imwbcnt == "8" | df$imwbcnt == "9" | df$imwbcnt == "10", 3, df$imwbcnt)
df$imwbcnt <- ifelse(df$imwbcnt == "4" | df$imwbcnt == "5" | df$imwbcnt == "6", 2, df$imwbcnt)
df$imwbcnt <- ifelse(df$imwbcnt == "0" | df$imwbcnt == "1" | df$imwbcnt == "2" | df$imwbcnt == "3", 1, df$imwbcnt)
```

```{r}
#demographics recodes```
df$agea <- ifelse(df$agea  == "999", NA, df$agea)

df$gndr <- ifelse(df$gndr  == "9", NA, df$gndr)
df$gndr <- ifelse(df$gndr  == "1", "Male", "Female")

df$edulvla <- ifelse(df$edulvla == "0" | df$edulvla == "77" | df$edulvla == "88" | df$edulvla == "99", NA, df$edulvla)
df$edulvla <- factor(df$edulvla,  levels=1:6)

df$ctzcntr <- ifelse(df$ctzcntr  == "7" | df$ctzcntr  == "8" | df$ctzcntr  == "9" , NA, df$ctzcntr)
df$ctzcntr <- factor(df$ctzcntr, labels=c('Citizen', 'Non-citizen'))
```

```{r}
df$stflife <- ifelse(df$stflife == "77" | df$stflife == "88" | df$stflife == "99", NA, df$stflife)
df$stflife <- ifelse(df$stflife == "7" | df$stflife == "8" | df$stflife == "9" | df$stflife == "10", 3, df$stflife)
df$stflife <- ifelse(df$stflife == "4" | df$stflife == "5" | df$stflife == "6", 2, df$stflife)
df$stflife <- ifelse(df$stflife == "0" | df$stflife == "1" | df$stflife == "2" | df$stflife == "3", 1, df$stflife)

#stfeco: how satisfied with present state of economy
df$stfeco <- ifelse(df$stfeco == "77" | df$stfeco == "88" | df$stfeco == "99", NA, df$stfeco)
df$stfeco <- ifelse(df$stfeco == "7" | df$stfeco == "8" | df$stfeco == "9" | df$stfeco == "10", 3, df$stfeco)
df$stfeco <- ifelse(df$stfeco == "4" | df$stfeco == "5" | df$stfeco == "6", 2, df$stfeco)
df$stfeco <- ifelse(df$stfeco == "0" | df$stfeco == "1" | df$stfeco == "2" | df$stfeco == "3", 1, df$stfeco)

#stfgov: how satisfied with national government
df$stfgov <- ifelse(df$stfgov == "77" | df$stfgov == "88" | df$stfgov == "99", NA, df$stfgov)
df$stfgov <- ifelse(df$stfgov == "7" | df$stfgov == "8" | df$stfgov == "9" | df$stfgov == "10", 3, df$stfgov)
df$stfgov <- ifelse(df$stfgov == "4" | df$stfgov == "5" | df$stfgov == "6", 2, df$stfgov)
df$stfgov <- ifelse(df$stfgov == "0" | df$stfgov == "1" | df$stfgov == "2" | df$stfgov == "3", 1, df$stfgov)

#stfdem: how satisfied with the way democracy works in  country
df$stfdem <- ifelse(df$stfdem == "77" | df$stfdem == "88" | df$stfdem == "99", NA, df$stfdem)
df$stfdem <- ifelse(df$stfdem == "7" | df$stfdem == "8" | df$stfdem == "9" | df$stfdem == "10", 3, df$stfdem)
df$stfdem <- ifelse(df$stfdem == "4" | df$stfdem == "5" | df$stfdem == "6", 2, df$stfdem)
df$stfdem <- ifelse(df$stfdem == "0" | df$stfdem == "1" | df$stfdem == "2" | df$stfdem == "3", 1, df$stfdem)


#happy: how happy are you
df$happy <- ifelse(df$happy == "77" | df$happy == "88" | df$happy == "99", NA, df$happy)
df$happy <- ifelse(df$happy == "0" | df$happy == "1" | df$happy == "2" | df$happy == "3" | df$happy == "4", 1, df$happy)
df$happy <- ifelse(df$happy == "5", 2, df$happy)
df$happy <- ifelse(df$happy == "6" | df$happy == "7" | df$happy == "8" | df$happy == "9" | df$happy == "10", 3, df$happy)

df$health <- ifelse(df$imsmetn  == "7" | df$health  == "8" | df$health  == "9", NA, df$health)
df$ipeqopt <- ifelse(df$ipeqopt  == "7" | df$ipeqopt  == "8" | df$ipeqopt  == "9", NA, df$ipeqopt)
df$ipudrst <- ifelse(df$ipudrst  == "7" | df$ipudrst  == "8" | df$ipudrst  == "9", NA, df$ipudrst)
df$ipbhprp <- ifelse(df$ipbhprp  == "7" | df$ipbhprp  == "8" | df$ipbhprp  == "9", NA, df$ipbhprp)
df$impenv <- ifelse(df$impenv  == "7" | df$impenv  == "8" | df$impenv  == "9", NA, df$impenv)
df$imptrad <- ifelse(df$imptrad  == "7" | df$imptrad  == "8" | df$imptrad  == "9", NA, df$imptrad)

df <- df %>%
  mutate(health = 6 - health,
         ipeqopt = 7 - ipeqopt,
         ipudrst = 7 - ipudrst,
         ipbhprp = 7 - ipbhprp,
         impenv = 7 - impenv, 
         imptrad = 7 - imptrad)

#health: subjective general health
df$health <- ifelse(df$health == "5"| df$health == "4", 1, df$health)
df$health <- ifelse(df$health == "3", 2, df$health)
df$health <- ifelse(df$health == "4" | df$health == "5", 3, df$health)

#ipeqopt: important that people are treated equally and have equal opportunities
df$ipeqopt <- ifelse(df$ipeqopt == "5" | df$ipeqopt == "6", 3, df$ipeqopt)
df$ipeqopt <- ifelse(df$ipeqopt == "3" | df$ipeqopt == "4", 2, df$ipeqopt)
df$ipeqopt <- ifelse(df$ipeqopt == "1" | df$ipeqopt == "2", 1, df$ipeqopt)

#ipudrst: important to understand different people
df$ipudrst <- ifelse(df$ipudrst == "5" | df$ipudrst == "6", 3, df$ipudrst)
df$ipudrst <- ifelse(df$ipudrst == "3" | df$ipudrst == "4" , 2, df$ipudrst)
df$ipudrst <- ifelse(df$ipudrst == "1" | df$ipudrst == "2", 1, df$ipudrst)

#ipbhprp: important to behave properly
df$ipbhprp <- ifelse(df$ipbhprp == "5" | df$ipbhprp == "6", 3, df$ipbhprp)
df$ipbhprp <- ifelse(df$ipbhprp == "3" | df$ipbhprp == "4" , 2, df$ipbhprp)
df$ipbhprp <- ifelse(df$ipbhprp == "1" | df$ipbhprp == "2", 1, df$ipbhprp)

#impenv: important to care for nature and environment
df$impenv <- ifelse(df$impenv == "5" | df$impenv == "6", 3, df$impenv)
df$impenv <- ifelse(df$impenv == "3" | df$impenv == "4" , 2, df$impenv)
df$impenv <- ifelse(df$impenv == "1" | df$impenv == "2", 1, df$impenv)

#imptrad: important to follow traditions and customs
df$imptrad <- ifelse(df$imptrad == "5" | df$imptrad == "6", 3, df$imptrad)
df$imptrad <- ifelse(df$imptrad == "3" | df$imptrad == "4", 2, df$imptrad)
df$imptrad <- ifelse(df$imptrad == "1" | df$imptrad == "2", 1, df$imptrad)
```


```{r}
df[is.na(df)] <- 0

df04  <- df %>%
  filter(essround==2)
df12  <- df %>%
  filter(essround==6)
df16  <- df %>%
  filter(essround==8)
df18  <- df %>%
  filter(essround==9)
```

```{r factor analysis by year }
#2004
summary(df04)

corr04 <- cor(df04[19:24])
n.obs <- nrow(df04)
round(corr04)
corr04 <- as.matrix(corr04)
cor.plot(corr04)

df04.1 <- df04
#df04.1[is.na(df04.1)] <- 0
df04.1t <- df04.1[-c(1:18, 25:80)]

corr04mod <- fa(corr04, nfactors=2, n.obs=n.obs, fm='minres', rotate='varimax')
fa.diagram(corr04mod)
corr04mod$loadings

scores04 <- factor.scores(df04.1t, corr04mod, method="Thurstone")
scoresdf04 <- scores04$scores
df04.1 <- cbind(df04.1, scoresdf04)
```

```{r FA 2012}
#2004
summary(df04)

corr04 <- cor(df04[19:24])
n.obs <- nrow(df04)
round(corr04)
corr04 <- as.matrix(corr04)
cor.plot(corr04)

df04.1 <- df04
#df04.1[is.na(df04.1)] <- 0
df04.1t <- df04.1[-c(1:18, 25:80)]

corr04mod <- fa(corr04, nfactors=2, n.obs=n.obs, fm='minres', rotate='varimax')
fa.diagram(corr04mod)
corr04mod$loadings

scores04 <- factor.scores(df04.1t, corr04mod, method="Thurstone")
scoresdf04 <- scores04$scores
df04.1 <- cbind(df04.1, scoresdf04)
```

```{r FA 2012}
#2012

corr12 <- cor(df12[19:24])
n.obs <- nrow(df12)
round(corr12)
corr12 <- as.matrix(corr12)
cor.plot(corr12)

df12.1 <- df12
#df12.1[is.na(df12.1)] <- 0
df12.1t <- df12.1[-c(1:18, 25:80)]

corr12mod <- fa(corr12, nfactors=2, n.obs=n.obs, fm='minres', rotate='varimax')
fa.diagram(corr12mod)
corr12mod$loadings

scores12 <- factor.scores(df12.1t, corr12mod, method="Thurstone")
scoresdf12 <- scores12$scores
df12.1 <- cbind(df12.1, scoresdf12)
```
```{r FA 2016}
#2016

corr16 <- cor(df16[19:24])
n.obs <- nrow(df16)
round(corr16)
corr16 <- as.matrix(corr16)
cor.plot(corr16)

df16.1 <- df16
#df16.1[is.na(df12.1)] <- 0
df16.1t <- df16.1[-c(1:18, 25:80)]

corr16mod <- fa(corr16, nfactors=2, n.obs=n.obs, fm='minres', rotate='varimax')
fa.diagram(corr16mod)
corr16mod$loadings

scores16 <- factor.scores(df16.1t, corr16mod, method="Thurstone")
scoresdf16 <- scores16$scores
df16.1 <- cbind(df16.1, scoresdf16)
```

```{r FA 2018}
#2018
corr18 <- cor(df18[19:24])
n.obs <- nrow(df18)
round(corr18)
corr18 <- as.matrix(corr18)
cor.plot(corr18)

df18.1 <- df18
#df16.1[is.na(df12.1)] <- 0
df18.1t <- df18.1[-c(1:18, 25:80)]

corr18mod <- fa(corr18, nfactors=2, n.obs=n.obs, fm='minres', rotate='varimax')
fa.diagram(corr18mod)
corr18mod$loadings

scores18 <- factor.scores(df18.1t, corr18mod, method="Thurstone")
scoresdf18 <- scores18$scores
df18.1 <- cbind(df18.1, scoresdf18)
```

```{r factor analysis}
corr1 <- cor(df[19:24])
n.obs <- nrow(df)
round(corr1)
corr1 <- as.matrix(corr1)
cor.plot(corr1)
#corr1mod <- fa(corr1, nfactors=2, n.obs=n.obs, fm='minres', rotate='varimax')
#fa.diagram(corr1mod)

df2 <- df
df2[is.na(df)] <- 0
df2t <- df2[-c(1:18, 25:80)]

corr1 <- cor(df2t)
corr1mod <- fa(corr1, nfactors=2, n.obs=n.obs, fm='minres', rotate='varimax')
fa.diagram(corr1mod)

corr1mod$loadings
#corr1mod$communalities
#corr1mod$uniquenesses
#corr1mod$complexity

scores1 <- factor.scores(df2t, corr1mod, method="Thurstone")
scoresdf <- scores1$scores
df2 <- cbind(df2, scoresdf)
head(scores1$scores, 10)

#df2 <- df
#df2[is.na(df)] <- 0

#df2$imdfetn <- -1 * df2t$imdfetn

#df2t <- df2[-c(1:18, 25:80)]

#corr2 <- cor(df2t)
#corr2mod <- fa(corr2, nfactors=2, n.obs=n.obs, fm='minres', rotate='varimax')
#fa.diagram(corr2mod)
#corr2mod$loadings

#corr2mod$communalities
#corr2mod$uniquenesses
#corr2mod$complexity
#summary(corr2mod)
#scores1 <- factor.scores(df2t, corr2mod, method="Thurstone")
#head(scores1$scores, 10)
#scoresdf <- scores1$scores
#df2 <- cbind(df2, scoresdf)
#head(df2)

alpha <- alpha(corr1)
alpha
```

```{r}

df2$essround <- ifelse(df2$essround == "2", 360400, df2$essround)
df2$essround <- ifelse(df2$essround == "6", 672900, df2$essround)
df2$essround <- ifelse(df2$essround == "8", 1792200, df2$essround)
df2$essround <- ifelse(df2$essround == "9", 2315925, df2$essround)
df2$essround <- factor(df2$essround)

regmod1 <- lm(MR1 ~ essround + happy + stflife + stfeco + stfgov + ipeqopt + ipudrst + imptrad, df2)
regmod2 <- lm(MR2 ~ essround + happy + stflife + stfeco  + stfgov + ipeqopt + ipudrst + imptrad, df2)

regmod12 <- lm(MR1 ~ essround, df2)
summary(regmod1)
regmod12 <- lm(MR2 ~ essround, df2)
summary(regmod1)

```
```{r}
mod1 <- regmod1 <- lm(MR1 ~ tourists, df2)

```

